### 2019-09-03 C11-BackEnds/Clean/document/root.tex
Brought documentation roughly into shape.

\documentclass[fontsize=11pt,paper=a4,open=right,twoside,abstract=true]{scrreprt}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[numbers, sort&compress, sectionbib]{natbib}
\usepackage{isabelle,isabellesym}
\usepackage[only,bigsqcap]{stmaryrd}
\usepackage{ifthen}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{lstisadof}
\IfFileExists{railsetup.sty}{\usepackage{railsetup}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% command

\newenvironment{matharray}[1]{\[\begin{array}{#1}}{\end{array}\]} % from 'iman.sty'
\newcommand{\indexdef}[3]%
{\ifthenelse{\equal{}{#1}}{\index{#3 (#2)|bold}}{\index{#3 (#1\ #2)|bold}}} % from 'isar.sty'
\newcommand{\isactrlC}{{\bf C}}

%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% fix for package declaration to be at the end
\usepackage[pdfpagelabels, pageanchor=false, plainpages=false]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% document

\urlstyle{rm}
\isabellestyle{it}

\newcommand{\HOL}{HOL}
\newcommand{\eg}{e.g.}
\newcommand{\ie}{i.e.}

\begin{document}

\title{Clean - An Abstract Imperative Programming Language and its Theory}
\author{Frederic Tuong and Burkhart Wolff \\ (with Contributions of Chantal Keller)}
\maketitle

\begin{abstract}
Clean (pronounced as: ``C lean'') is based on a simple, abstract execution model for an 
imperative target language. "Abstract" is understood as contrast to "Concrete Semantics";
alternatively, the term "shallow-style embedding" could be used.

Clean is based on a ``no-frills'' state-exception monad  with the 
usual definitions of \verb+bind+ and \verb+unit+ for the compositional glue of
state-based computations. Clean offers conditionals and loops supporting C-like 
control-flow operators such as \verb+break+ and \verb+return+.
The memory model is fully \HOL -typed and based on a construction nesting
subsequently extensible records. Note that the current version only supports
monomorphic states and parameters of (directly recursive) functions.

The design is striving for extreme simplicity. It is geared towards symbolic execution and
proven correct verification tools. The underlying libraries of this package, however,
deliberately restricts itself to the most elementary infrastructure for these tasks. 
The package is intended to serve as demonstrator semantic backend for Isabelle/C \cite{TuongWolff19},
serving as basis for the test-generation technique described in \cite{DBLP:conf/tap/Keller18} .
\end{abstract}

\newpage
\tableofcontents

\parindent 0pt\parskip 0.5ex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \input{session}
\newpage
\input{Clean.tex}

\input{Quicksort_concept.tex}

\newpage
\section{Appendix : Used Monad Libraries}
\input{MonadSE.tex}

\input{Seq_MonadSE.tex}

\input{Symbex_MonadSE.tex}

\input{SquareRoot_concept.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}
\bibliography{root}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: