### 2019-10-14 C11-FrontEnd/document/preamble.tex

%e% This is a placeholder for user-specific configuration and packages.
\usepackage{lmodern}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{paralist}
\usepackage[en-US]{datetime2}      % For \DTMdate, etc.
\usepackage{amssymb}
\usepackage{listings}
\usepackage{lstisadof}
\usepackage{xspace}
\usepackage[draft]{fixme}
\usepackage{wrapfig}

\title{Deeply Integrating C11 Code Support into Isabelle/PIDE} % should be auto-generated
\addauthor{Frédéric Tuong\inst {1}} % should be auto-generated

\newcommand{\house}{\mathrel{\substack{\wedge\\[-.2em]\sqcup}}}