### 2019-10-15 C11-FrontEnd/document/preamble.tex
synchronize with fortesse revision 5366

%e% This is a placeholder for user-specific configuration and packages.
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{paralist}
\usepackage[en-US]{datetime2}      % For \DTMdate, etc.
\usepackage{amssymb}
\usepackage{listings}
\usepackage{lstisadof}
\usepackage{xspace}
\usepackage[draft]{fixme}
\usepackage{wrapfig}

\title{} % should be auto-generated
\addauthor{\href{https://www.lri.fr/~ftuong/}{Frédéric Tuong}\inst{1} \and \href{https://www.lri.fr/~wolff/}{Burkhart Wolff}\inst{1}} % should be auto-generated

\newcommand{\house}{\mathrel{\substack{\wedge\\[-.2em]\sqcup}}}