### 2021-12-28 Citadelle/src/uml_main/document/root.tex

\documentclass[fontsize=10pt,DIV12,paper=a4,open=right,twoside,abstract=true]{scrreprt}
\usepackage{fixltx2e}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[english]{babel}
\usepackage{isabelle}
\isatagannexa
  \usepackage{omg}
  \usepackage{draftwatermark}
  \SetWatermarkAngle{55}
  \SetWatermarkLightness{.9}
  \SetWatermarkFontSize{3cm}
  \SetWatermarkScale{1.4}
  \SetWatermarkText{\textbf{\textsf{Draft Proposal}}}
\endisatagannexa
\usepackage[alpine]{ifsym}
\usepackage[nocolortable, noaclist,isasymonly,nocolor]{hol-ocl-isar}
\renewcommand{\lfloor}{\isasymHolOclLiftLeft}
\renewcommand{\rfloor}{\isasymHolOclLiftRight}
\renewcommand{\lceil}{\isasymHolOclDropLeft}
\renewcommand{\rceil}{\isasymHolOclDropRight}
\renewcommand{\oclkeywordstyle}{\bfseries}
\renewcommand{\javakeywordstyle}{\bfseries}
\renewcommand{\smlkeywordstyle}{\bfseries}
\renewcommand{\holoclthykeywordstyle}{}
\renewcommand{\greenkeywordstyle}{}

\usepackage{lstisar}
\usepackage{railsetup}
\usepackage[]{mathtools}
\usepackage{%
  multirow,
  paralist,
  booktabs,     %       "        "      "
  threeparttable,
  longtable,    % Mehrseitige Tabellen 
}

\usepackage{graphicx}
\usepackage[numbers, sort&compress, sectionbib]{natbib}
\usepackage{chapterbib}
\usepackage[caption=false]{subfig}
\usepackage{tabu}
\usepackage{prooftree}
\usepackage[draft]{fixme}
\usepackage[pdfpagelabels, pageanchor=false, bookmarksnumbered, plainpages=false]{hyperref}
\graphicspath{{data/},{figures/}}
\makeatletter
\renewcommand*\l@section{\bprot@dottedtocline{1}{1.5em}{2.8em}}
\renewcommand*\l@subsection{\bprot@dottedtocline{2}{3.8em}{3.7em}}
\renewcommand*\l@subsubsection{\bprot@dottedtocline{3}{7.0em}{5em}}
\renewcommand*\l@paragraph{\bprot@dottedtocline{4}{10em}{6.2em}}
%\renewcommand*\l@paragraph{\bprot@dottedtocline{4}{10em}{5.5em}}
\renewcommand*\l@subparagraph{\bprot@dottedtocline{5}{12em}{7.7em}}
%\renewcommand*\l@subparagraph{\bprot@dottedtocline{5}{12em}{6.5em}}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Overall the (rightfully issued) warning by Koma Script that \rm
%%% etc. should not be used (they are deprecated since more than a
%%% decade)
  \DeclareOldFontCommand{\rm}{\normalfont\rmfamily}{\mathrm}
  \DeclareOldFontCommand{\sf}{\normalfont\sffamily}{\mathsf}
  \DeclareOldFontCommand{\tt}{\normalfont\ttfamily}{\mathtt}
  \DeclareOldFontCommand{\bf}{\normalfont\bfseries}{\mathbf}
  \DeclareOldFontCommand{\it}{\normalfont\itshape}{\mathit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{tocdepth}{3} % printed TOC not too detailed
\hypersetup{bookmarksdepth=3} % more detailed digital TOC (aka bookmarks)
\sloppy
\allowdisplaybreaks[4]
\raggedbottom

\newcommand{\FOL}{FOL\xspace}
\newcommand{\HOL}{HOL\xspace}
\newcommand{\OCL}{OCL\xspace}
\newcommand{\UML}{UML\xspace}
\newcommand{\HOLOCL}{HOL-OCL\xspace}
\newcommand{\FOCL}{Featherweight OCL\xspace}
\newcommand{\ML}{ML\xspace}
\newcommand{\HOCL}{HOL-OCL 2.0\xspace}
\renewcommand{\HolTrue}{\mathrm{True}}
\renewcommand{\HolFalse}{\mathrm{False}}
\newcommand{\ptmi}[1]{\using{\mi{#1}}}
\newcommand{\Lemma}[1]{{\color{BrickRed}%
    \mathbf{\operatorname{lemma}}}~\text{#1:}\quad}
\newcommand{\done}{{\color{OliveGreen}\operatorname{done}}}
\newcommand{\apply}[1]{{\holoclthykeywordstyle%
    \operatorname{apply}}(\text{#1})}
\newcommand{\IIbegin} {{\operatorname{begin}}}
\newcommand{\IIend} {{\operatorname{end}}}
\newcommand{\fun} {{\holoclthykeywordstyle\operatorname{fun}}}
\newcommand{\overloading} {{\holoclthykeywordstyle\operatorname{overloading}}}
\newcommand{\isardef} {{\holoclthykeywordstyle\operatorname{definition}}}
\newcommand{\where} {{\holoclthykeywordstyle\operatorname{where}}}
\newcommand{\datatype} {{\holoclthykeywordstyle\operatorname{datatype}}}
\newcommand{\types} {{\holoclthykeywordstyle\operatorname{types}}}
\newcommand{\Instance} {{\holoclthykeywordstyle\operatorname{Instance}}}
\newcommand{\State} {{\holoclthykeywordstyle\operatorname{State}}}
\newcommand{\Transition} {{\holoclthykeywordstyle\operatorname{Transition}}}
\newcommand{\record} {{\holoclthykeywordstyle\operatorname{record}}}
\newcommand{\tysequence} {\text{\ttfamily Sequence}}
\newcommand{\pglabel}[1]{\text{#1}}
\renewcommand{\isasymOclUndefined}{\ensuremath{\mathtt{invalid}}}
\newcommand{\isasymOclNull}{\ensuremath{\mathtt{null}}}
\newcommand{\isasymOclInvalid}{\isasymOclUndefined}
\DeclareMathOperator{\inv}{inv}
\newcommand{\Null}[1]{{\ensuremath{\mathtt{null}_\text{{#1}}}}}
\newcommand{\testgen}{HOL-TestGen\xspace}
\newcommand{\HolOption}{\mathrm{option}}
\renewcommand{\to}{\Rightarrow}
\newcommand{\ran}{\mathrm{ran}}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\typedef}{\mathrm{typedef}}
\newcommand{\typesynonym}{\mathrm{type\_synonym}}
\newcommand{\mi}[1]{\,\text{#1}}
\newcommand{\state}[1]{(#1) \operatorname{state}}
\newcommand{\mocl}[1]{\text{\inlineocl|#1|}}
\DeclareMathOperator{\TCnull}{null}
\DeclareMathOperator{\HolNull}{null}
\DeclareMathOperator{\HolBot}{bot}
\newcommand{\isaAA}{\mathfrak{A}}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}
\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.\xspace}

\newenvironment{isamarkuplazy_text}{\par \isacommand{lazy{\isacharunderscore}text}\isamarkupfalse\isacharverbatimopen\isastyletext\begin{isapar}}{\end{isapar}\isacharverbatimclose}
\newcommand{\isactrlmkUNDERSCOREstring}{\isakeywordcontrol{mk{\isacharunderscore}string}}
\renewcommand{\isasymguillemotleft}{\isatext{\textquotedblleft}}
\renewcommand{\isasymguillemotright}{\isatext{\textquotedblright}}
\begin{document}
\renewcommand{\subsubsectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section}
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\chapterautorefname}{Chapter}
\newcommand{\subtableautorefname}{\tableautorefname}
\newcommand{\subfigureautorefname}{\figureautorefname}
\isatagannexa
\renewcommand\thepart{\Alph{part}}
\renewcommand\partname{Annex}
\endisatagannexa

\newenvironment{defholsimple}{\textbf{Definition}\xspace}{}

\newenvironment{matharray}[1]{\[\begin{array}{#1}}{\end{array}\]} % from 'iman.sty'
\newcommand{\indexdef}[3]%
{\ifthenelse{\equal{}{#1}}{\index{#3 (#2)|bold}}{\index{#3 (#1\ #2)|bold}}} % from 'isar.sty'

\isatagafp
  \title{Featherweight OCL}
  \subtitle{A Proposal for a Machine-Checked Formal Semantics for OCL 2.5 %\\
    %\includegraphics[scale=.5]{figures/logo_focl}
  }
\endisatagafp
\isatagannexa
  \title{A Formal Machine-Checked Semantics for OCL 2.5}
  \subtitle{A Proposal for the "Annex A" of the OCL Standard}
\endisatagannexa
\isatagnoexample
  \title{Featherweight OCL (without ``employee'' examples)}
\endisatagnoexample
\author{%
  \href{http://www.brucker.ch/}{Achim D. Brucker}\footnotemark[1]
  \and
  \href{https://www.lri.fr/~tuong/}{Fr\'ed\'eric Tuong}\footnotemark[2]~\footnotemark[3]
  \and
  \href{https://www.lri.fr/~wolff/}{Burkhart Wolff}\footnotemark[2]~\footnotemark[3]}
\publishers{%
  \footnotemark[1]~SAP SE\\
  Vincenz-Priessnitz-Str. 1, 76131 Karlsruhe,
  Germany \texorpdfstring{\\}{} \href{mailto:"Achim D. Brucker"
    <achim.brucker@sap.com>}{achim.brucker@sap.com}\\[2em]
  %
  \footnotemark[2]~LRI, Univ. Paris-Sud, CNRS, CentraleSup\'elec, Universit\'e Paris-Saclay \\
  b\^at. 650 Ada Lovelace, 91405 Orsay, France \texorpdfstring{\\}{}
    \href{mailto:"Frederic Tuong"
    <frederic.tuong@lri.fr>}{frederic.tuong@lri.fr} \hspace{4.5em}
    \href{mailto:"Burkhart Wolff"
    <burkhart.wolff@lri.fr>}{burkhart.wolff@lri.fr} \\[2em]
  %
  \footnotemark[3]~IRT SystemX\\
  8 av.~de la Vauve, 91120 Palaiseau, France \texorpdfstring{\\}{}
    \href{mailto:"Frederic Tuong"
    <frederic.tuong@irt-systemx.fr>}{frederic.tuong@irt-systemx.fr} \quad
    \href{mailto:"Burkhart Wolff"
    <burkhart.wolff@irt-systemx.fr>}{burkhart.wolff@irt-systemx.fr}
}

\maketitle
\isatagannexa
\cleardoublepage
\endisatagannexa

\isatagafp
  \begin{abstract}
    The Unified Modeling Language (UML) is one of the few modeling
    languages that is widely used in industry. While UML is mostly known
    as diagrammatic modeling language (\eg, visualizing class models),
    it is complemented by a textual language, called Object Constraint
    Language (OCL). OCL is a textual annotation language, originally based on a
    three-valued logic, that turns UML into a formal language.
    Unfortunately the semantics of this specification language, captured
    in the ``Annex A'' of the OCL standard, leads to different
    interpretations of corner cases.  Many of these corner cases had
    been subject to formal analysis since more than ten years.

    The situation complicated with the arrival of version 2.3 of the OCL 
    standard. OCL was aligned with the latest version of UML: this led to the 
    extension of the three-valued logic by a second exception element, called
    \inlineocl{null}.  While the first exception element
    \inlineocl{invalid} has a strict semantics, \inlineocl{null} has a
    non strict interpretation. The combination of these semantic features lead
    to remarkable confusion for implementors of OCL compilers and
    interpreters.

    In this paper, we provide a formalization of the core of OCL in
    HOL\@. It provides denotational definitions, a logical calculus and
    operational rules that allow for the execution of OCL expressions by
    a mixture of term rewriting and code compilation. Moreover, we describe
    a coding-scheme for UML class models that were annotated by 
    code-invariants and code contracts. An implementation of this coding-scheme
    has been undertaken: it consists of a kind of compiler that takes a UML class
    model and translates it into a family of definitions and derived
    theorems over them capturing the properties of constructors and selectors,
    tests and casts resulting from the class model. However, this compiler
    is \emph{not} included in this document.

    Our formalization reveals several inconsistencies and contradictions
    in the current version of the OCL standard.  They reflect a challenge
    to define and implement OCL tools in a uniform manner.  Overall, this
    document is intended to provide the basis for a machine-checked text
    ``Annex A'' of the OCL standard targeting at tool implementors.

%% Le Langage de Modélisation Unifié (UML) fait partie des quelques
%% langages de modélisation couramment utilisés dans l'industrie. Alors
%% qu'UML est bien souvent connu comme langage de modélisation de
%% diagrammes (par exemple, comme classes de modèles visuels), il est
%% plus précisément complémenté par un langage textuel, appelé Langage
%% Contraint d'Objet (OCL). OCL est un langage d'annotation textuel,
%% originellement basé sur une logique 3-valuée, et fait ainsi apparaître
%% UML comme un langage plutôt formel. Pourtant la sémantique de ce
%% langage de spécification, instinctivement décrit dans l'annexe A du
%% standard d'OCL, mène à différentes interprétations concernant le sens
%% précis des informations véhiculées. Ces différentes interprétations
%% sont encore en cours d'élucidation depuis plus de dix ans via
%% notamment l'emploi d'outils d'analyse formelle.
%% 
%% La situation s'est compliquée avec la mise à jour vers la version 2.3
%% du standard d'OCL. Étant donné qu'OCL dépends de la dernière version
%% d'UML, il a été nécessaire d'introduire comme extension à la logique
%% 3-valuée un second élément d'exception, qui est null. Ce qui le
%% différencie avec le premier élément d'exception invalid, c'est sa
%% sémantique non-stricte, alors qu'invalid a une sémantique stricte. La
%% combinaison de ces variétés sémantiques a finalement abouti à divers
%% confusions lors de l'implémentation de compilateurs OCL.
%% 
%% Dans ce papier, nous apportons une formalisation du cœur d'OCL en
%% HOL. Il contient des définitions dénotationnelles, un système de
%% calcul logique et des règles opérationnelles permettant l'exécution
%% des expressions OCL à travers une mixture de règles de réécriture et
%% de compilation de code. De plus, nous décrivons un schéma de codage
%% pour les modèles de classes UML pour représenter les invariants et
%% contrats UML. Une implémentation de ce schéma de codage a été
%% effectuée : il s'agit d'un genre particulier de compilateur acceptant
%% en entrée des classes de modèles UML et générant en sortie une famille
%% de définitions, théorèmes associés sur les propriétés des
%% constructeurs et sélecteurs, fonctions de tests ainsi que des
%% mutations de types. Ce compilateur n'est cependant pas inclus dans ce
%% document.
%% 
%% Nos formalisations ont finalement révélés des inconsistances et
%% contradictions concernant la version courante du standard d'OCL. Par
%% ailleurs cela représente un challenge de définir et implémenter des
%% outils pour OCL de manière uniformément cohérente. Ce document a pour
%% but de fournir des bases vérifiées mécaniquement, et servir comme
%% support sémantique à l'Annexe A du standard d'OCL ciblant
%% particulièrement les programmeurs d'outils et compilateurs.
  \end{abstract}
  \tableofcontents
\endisatagafp

\part{Formal Semantics of OCL}
\input{introduction}
%\clearpage
\isatagafp
\input{session}
\endisatagafp
\isatagnoexample
\input{session}
\endisatagnoexample
\isatagannexa
\input{UML_Types.tex}
\input{UML_Logic.tex}
\input{UML_PropertyProfiles.tex}
\input{UML_Boolean.tex}
\input{UML_Void.tex}
\input{UML_Integer.tex}
\input{UML_Real.tex}
\input{UML_String.tex}
\input{UML_Pair.tex}
\input{UML_Bag.tex}
\input{UML_Set.tex}
\input{UML_Sequence.tex}
\input{UML_Library.tex}
\input{UML_State.tex}
\input{UML_Contracts.tex}
%\input{UML_Tools.tex}
%\input{UML_Main.tex}
% \input{Design_UML.tex}
% \input{Design_OCL.tex}
\input{Analysis_UML.tex}
\input{Analysis_OCL.tex}
\part{Bibliography}
\endisatagannexa
\isatagafp
\input{conclusion} %no conclusion for standard document
\endisatagafp
\bibliographystyle{abbrvnat}
\bibliography{root}

\isatagafp
\appendix
\part{Appendix}
\endisatagafp
\input{FOCL_Syntax}

\isatagannexa
  \part{Table of Contents}
  \clearpage {\small \tableofcontents }
\endisatagannexa
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%  LocalWords:  implementors denotational OCL UML